The Pair Distribution Function of the Inhomogeneous Hard-Sphere Fluid

### 2014-02-25 papers/pair-correlation/paper.tex
clean up pairs text Tested on bennet. Took 5:27.21.

We construct $g(r, g_\sigma)$ to satisfy several constraints.  Naturally,
\begin{equation}
  g(\sigma, g_\sigma) = g_\sigma.
\end{equation}
A second constraint applies to the integral of the total correlation
function
\begin{align}
  1 + n\int h(r)d\rr &= nkT\chi_T \label{eq:total-constraint}
\end{align}
where $h(r) = g(r) - 1$ is the total correlation function.
A final constraint is that the slope of the radial distribution
function at contact is given to a very good approximation by \fixme{fix dimensions below}
\begin{align}
  g'(\sigma) \approx - h_\sigma g_\sigma
\end{align}
where $h_\sigma = g_\sigma - 1$ is the total correlation function.
Under these three constraints, we applied a least-squares fit to
simulation data for the radial distribution function at packing
fractions from 0.05 to 0.4 by steps of 0.05.  Our fitted radial
distribution function is given by
\newcommand\kappaa{\kappa_0}
\newcommand\kappab{\kappa_1}
\newcommand\kappac{\kappa_2}
\begin{multline}
  h(r,g_\sigma) = 1 + h_\sigma e^{-\kappaa z} \\
  + h_\sigma(\kappaa - g_\sigma) ze^{-\kappab z} \\
  + \frac{\frac{(1-\eta)^4}{1+4\eta + 4\eta^2 - 4\eta^3 +
      \eta^4}\frac{1}{n}
    - I_0  -I_1(\kappaa - g_\sigma)}{I_2} z^2e^{-\kappac z}
\end{multline}
\begin{equation}
  I_0 = \frac{4\pi h_\sigma (2 + \sigma \kappaa (2 + \sigma \kappaa))}{\kappaa^3}
  - \frac{4\pi}{3} \sigma^3
\end{equation}
\begin{equation}
  I_1 = \frac{4\pi h_\sigma (6 + \sigma \kappab (4 + \sigma \kappab))}{\kappab^4}
\end{equation}
\begin{equation}
  I_2 = \frac{8\pi (12 + \sigma \kappac (6 + \sigma \kappac))}{\kappac^5}
\end{equation}
\begin{equation}
  z = r - \sigma
\end{equation}
\davidsays{Can someone check the above equations for correctness
  against the python code, in plot-ghs2.py.  We'll want to
  ensure the C++ code agrees with this as well.}
\fixme{Clean up the above equations, if possible.}
This function includes three fitted parameters, $\kappaa$, $\kappab$,
and $\kappac$, which are given in Table~\ref{tab:fit-parameters}.
The fit is displayed in Fig.~\ref{fig:radial-distribution}.  In that
figure, it is apparent that the integral f $h(r)$ from
Eq.~\ref{eq:total-constraint} results in cancelling errors at large
distances.  The maximum error in $g(r)$ at packing fractions up to
0.4 is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
\rmaxerr$.

\begin{table}
  \begin{tabular}{cc}
    $\kappaa$ & \kappazero \\
    $\kappab$ & \kappaone \\
    $\kappac$ & \kappatwo
  \end{tabular}
  \caption{Fit parameters \fixme{... put them here and format this nicely}\fixme{add dimensions here}}
  \label{tab:fit-parameters}
\end{table}

%% In addition, we know that
%% \begin{align}
%%   1 + n\int\left[g(r)-1\right]d\rr &= nkT\chi_T
%% \end{align}
%% \begin{equation}
%%   \begin{split}
%%     \chi_T  &= -\frac{1}{V}\left(\frac{\partial V}{\partial
%%       p}\right)_T \\
%%     &= - \left(V\left(\frac{\partial p}{\partial
%%       V}\right)_T\right)^{-1}
%%   \end{split}
%% \end{equation}
%% \begin{equation}
%%     p = nkT\frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}
%% \end{equation}
%% \begin{equation}
%%     \left(\frac{\partial p}{\partial V}\right)_T = \frac{\partial p}{\partial\eta}\frac{\partial\eta}{\partial V} + \frac{\partial p}{\partial n}\frac{\partial n}{\partial V}
%% \end{equation}
%% \begin{equation}
%%     n = \frac{N}{V}\Longrightarrow \frac{\partial n}{\partial V} = -\frac{N}{V^2} = -\frac{n}{V}
%% \end{equation}
%% \begin{equation}
%%     \eta = \frac{4\pi R^3N}{3V}\Longrightarrow \frac{d\eta}{dV} = -\frac{\eta}{V}
%% \end{equation}
%% \begin{equation}
%%   \begin{split}
%%     \left(\frac{\partial p}{\partial V}\right)_T &= -nkT\frac{\eta}{V}\left[\frac{1 + 2\eta - 3\eta^2}{(1-\eta)^3} + 3\frac{1 + \eta + \eta^2 - \eta^3}{(1-\eta)^4}\right]\\
%%     &\quad  - \frac{nkT}{V}\left[\frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}\right]\\
%%     &= -nkT\frac{\eta}{V}\left[\frac{(1+2\eta-3\eta^2)(1-\eta) + 3 + 3\eta + 3\eta^2 - 3\eta^3}{(1-\eta)^4}\right]\\
%%     &\quad  - \frac{nkT}{V}\left[\frac{(1+\eta+\eta^2-\eta^3)(1-\eta)}{(1-\eta)^4}\right]\\
%%     &= -\frac{nkT}{V}\left[\frac{4\eta + 4\eta^2 - 2\eta^3}{(1-\eta)^4} + \frac{1 - 2\eta^3 + \eta^4}{(1-\eta)^4}\right]\\
%%   \end{split}
%% \end{equation}
%% \begin{equation}
%%     nkT\chi_T = \frac{(1-\eta)^4}{1 + 4\eta + 4\eta^2 - 4\eta^3 + \eta^4}
%%   %\int (g(r)-1)d\rr &= \frac{\chi_T}{kT} \\
%%   %&= \eta \eta \cdots \fixme{fill in, double-check}
%% \end{equation}
%% where $\chi_T$ is the isothermal compressibility, for which we use the
%% Carnahan-Starling equation of state.

%% An additional constraint that we use sets the slope of $g(r,
%% g_\sigma)$ at $r=\sigma$.  \fixme{explain this}

%% Given these constraints, we use a functional form of
%% \begin{equation}
%%   g(r,g_\sigma) = 1 + A e^{-a(r-\sigma)}
%%   + B (r-\sigma) e^{-b(r-\sigma)} + C (r-\sigma)^2 e^{-c(r-\sigma)}
%% \end{equation}
%% where $A$, $B$, and $C$ are amplitudes determined by the above
%% constraints, and $a$, $b$, and $c$ are parameters fit to Monte Carlo
%% simulations of the hard-sphere radial distribution function.

@@

\begin{figure*}
  \begin{subfigure}{\textwidth}
    \includegraphics[width=\linewidth]{figs/pair-correlation-pretty-1.pdf}
    \vspace{-0.6cm}
    \caption{Pair distribution function near a hard wall with $\eta = 0.1$.}
    \label{fig:pair-distribution-1}
  \end{subfigure}
  \begin{subfigure}{\textwidth}
    \includegraphics[width=\linewidth]{figs/pair-correlation-pretty-3.pdf}
    \vspace{-0.6cm}
    \caption{Pair distribution function near a hard wall with $\eta = 0.3$.}
    \label{fig:pair-distribution-3}
  \end{subfigure}
  \caption{ \paircaption \fixme{Do we want to label the y-axis for the
      1d plots? } \davidsays{I think so...}
  }\label{fig:pair-distribution}
\end{figure*}

@@

\begin{figure*}
  \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-contact-3.pdf}\\
  \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
  \caption{\fixme{Tweak spacing, add a, b, c, etc. The 1d plots still
      have two curves for this work, do we want to remove one?
      \davidsays{I'm not sure, I might be tempted to keep both,
        although if we do we should explain the difference between
        them. (left-right mirror image)}} The triplet distribution function
    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
    $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
    shows the triplet distribution function, with white corresponding
    to the contact value of the \emph{pair} distribution function.
    The plot on the right shows the value of $g^{(3)}$ along the path
    shown as a dotted line on the plot to the
    left.}\label{fig:triplet-contact-distribution-3}
\end{figure*}
% \begin{figure}
%   \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-contact-4.pdf}
%   \caption{The triplet distribution function
%     $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.4, plotted when
%     $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
%     shows the triplet distribution function, with white corresponding
%     to the contact value of the \emph{pair} distribution function.
%     The plot on the right shows the value of $g^{(3)}$ along the path
%     shown as a dotted line on the plot to the
%     left.}\label{fig:triplet-contact-distribution-4}
% \end{figure}

@@

% \begin{figure}
%   \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
%   \caption{The triplet distribution function
%     $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
%     separated by the diameter of one sphere.}\label{fig:triplet-inbetween-distribution-3}
% \end{figure}
% \begin{figure}
%   \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-inbetween-4.pdf}
%   \caption{The triplet distribution function
%     $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
%     separated by the diameter of one sphere.}\label{fig:triplet-inbetween-distribution-4}
% \end{figure}

\fixme{The following paragraph needs to be rewritten when we've
  reorganized the figures. (Obviously.)}

Figures~\ref{fig:triplet-contact-distribution-3}-\ref{fig:triplet-inbetween-distribution-4}
compare the triplet distribution function computed using our
approximation to $g^{(2)}$ with results from Monte Carlo simulation.
In each figure, $\rr_1$ and $\rr_2$ are held fixed, while the third
position $\rr_3$ is varied.  The test-particle position in each case
is $\rr_1$, which is on the left-hand side of the figure.  In the
color plot on the left, the first and second spheres are displayed in
grey.  On the bottom half of figures \fixme{b,d,etc?} is the triplet
correlation function as computed using the approximation presented in
this work.  The top half of those figures displays the triplet
correlation function from Monte Carlo.  Figures~\fixme{a,c,e?} show
the triplet distribution function along the paths illustrated in the
corresponding 2D plots.

Figures~\fixme{which} and~\fixme{which} show the triplet distribution
function when two of the positions $\rr_1$ and $\rr_2$ are in contact.
The triplet distribution function predicted by this work is slightly
asymmetric, indicating a failure in thermodynamic consistency.  Once
again, we observe the lack of oscillations at large distances.
Figures~\fixme{which} and~\fixme{which} show the triplet distribution
function when two of the positions $\rr_1$ and $\rr_2$ are spaced so
that the third could just fit between them.  Again, the triplet
distribution function predicted by this work is slightly asymmetric,
and lacks the oscillations seen in the Monte Carlo simulations at
large distances.  In both cases, our agreement with simulation is
comparable to that of previous approximations.

@@

\begin{figure}
  \begin{subfigure}{\columnwidth}
    \includegraphics[width=\columnwidth]{figs/dadz-3-2.pdf}
    \vspace{-0.8cm}
    \caption{Sticky hard-sphere fluid}\label{fig:dadz-delta}
  \end{subfigure}
  \begin{subfigure}{\columnwidth}
    \includegraphics[width=\columnwidth]{figs/dadz-square-well-3.pdf}
    \vspace{-0.8cm}
    \caption{Hard-core square well fluid}\label{fig:dadz-square-well}
  \end{subfigure}
  \begin{subfigure}{\columnwidth}
    \includegraphics[width=\columnwidth]{figs/dadz-inverse-sixth-3.pdf}
    \vspace{-0.8cm}
    \caption{Hard-core inverse-sixth fluid}\label{fig:dadz-inverse-sixth}
  \end{subfigure}
  \caption{Plot of $\frac{da_1}{dz}$ near a hard wall.\fixme{Do we
      want y-axis ticks or no? Also, why does the data start at $z =
      0$? \davidsays{I think ticks on the vertical axis aren't needed
        because we are not including the scale (which depends on
        prefactors).  I think data starting at $z=0$ is a result of a
        (poor?) choice of where to define $z=0$.  It's the same
        definition we use in Fig. 2, so I think the two should stay
        consistent...}}}
  \label{fig:dadz}
\end{figure}

>>>>

We construct $g(r, g_\sigma)$ to satisfy several constraints.  Naturally,
\begin{equation}
  g(\sigma, g_\sigma) = g_\sigma.
\end{equation}
A second constraint applies to the integral of the total correlation
function
\begin{align}
  1 + n\int h(r)d\rr &= nkT\chi_T \label{eq:total-constraint}
\end{align}
where $h(r) = g(r) - 1$ is the total correlation function.
A final constraint is that the slope of the radial distribution
function at contact is given to a very good approximation by
\begin{align}
  g'(\sigma) \approx - h_\sigma g_\sigma
\end{align}
where $h_\sigma = g_\sigma - 1$ is the total correlation function.
Under these three constraints, we applied a least-squares fit to
simulation data for the radial distribution function at packing
fractions from 0.05 to 0.4 by steps of 0.05.  Our fitted radial
distribution function is given by
\newcommand\kappaO{\kappa_0}
\newcommand\kappaI{\kappa_1}
\newcommand\kappaZ{\kappa_2}
\begin{multline}
  g(r,g_\sigma) = 1 + h_\sigma e^{-\kappaO \left(\frac{r}{\sigma}-1\right)} \\
  + h_\sigma(\kappaO - 2g_\sigma)  \left(\frac{r}{\sigma}-1\right)e^{-\kappaI  \left(\frac{r}{\sigma}-1\right)} \\
  + I(g_\sigma)  \left(\frac{r}{\sigma}-1\right)^2e^{-\kappaZ  \left(\frac{r}{\sigma}-1\right)}
\end{multline}
\begin{equation}
  I(g_\sigma) = \kappaZ^5 \frac{
    \frac{\chi-1}{24\eta} - h_\sigma \frac{(\kappaO-2
      g_\sigma)(\kappaI^2 + 4 \kappaI + 6)}{\kappaI^4}
    - h_\sigma\frac{\kappaO^2 + 2\kappaO + 2}{\kappaO}
  }{2 \kappaZ^2 + 12 \kappaZ + 24}
\end{equation}
\begin{equation}
  \chi = \frac{(1-\eta)^4}{\eta^4 - 4\eta^3 + 4\eta^2 + 4\eta + 1}
\end{equation}
\newcommand\nastyetacuberoot{\sqrt[3]{54 g_\sigma^2 -
    6\sqrt{81g_\sigma^4 - 6g_\sigma^3}}}
\begin{equation}
  \eta = 1 - \frac{1}{\nastyetacuberoot} - \frac{\nastyetacuberoot}{6g_\sigma}
\end{equation}
This function includes three fitted parameters, $\kappaO =
\kappazero$, $\kappaI = \kappaone$,
and $\kappaZ = \kappatwo$.
The fit is displayed in Fig.~\ref{fig:radial-distribution}.  In that
figure, it is apparent that the integral f $h(r)$ from
Eq.~\ref{eq:total-constraint} results in cancelling errors at large
distances.  The maximum error in $g(r)$ at packing fractions up to
0.4 is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
\rmaxerr$.

@@

\begin{figure*}
  \begin{subfigure}{\textwidth}
    \includegraphics[width=\linewidth]{figs/pair-correlation-pretty-1.pdf}
    \vspace{-0.6cm}
    \caption{Pair distribution function near a hard wall with $\eta = 0.1$.}
    \label{fig:pair-distribution-1}
  \end{subfigure}
  \begin{subfigure}{\textwidth}
    \includegraphics[width=\linewidth]{figs/pair-correlation-pretty-3.pdf}
    \vspace{-0.6cm}
    \caption{Pair distribution function near a hard wall with $\eta = 0.3$.}
    \label{fig:pair-distribution-3}
  \end{subfigure}
  \caption{ \paircaption \fixme{Do we want to label the y-axis for the
      1d plots? } \davidsays{I think so... also add a) b) c) d)}
  }\label{fig:pair-distribution}
\end{figure*}

@@

\begin{figure*}
  \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-contact-3.pdf}\\
  \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
  \caption{\fixme{Tweak spacing, add a, b, c, etc. The 1d plots still
      have two curves for this work, do we want to remove one?
      \davidsays{I'm not sure, I might be tempted to keep both,
        although if we do we should explain the difference between
        them. (left-right mirror image)}} The triplet distribution function
    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
    $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
    shows the triplet distribution function, with white corresponding
    to the contact value of the \emph{pair} distribution function.
    The plot on the right shows the value of $g^{(3)}$ along the path
    shown as a dotted line on the plot to the
    left.}\label{fig:triplet-contact-distribution}
\end{figure*}

@@

Figure~\ref{fig:triplet-contact-distribution} compares the triplet
distribution function at a packing fraction of 0.3 computed using our
approximation to $g^{(2)}$ with results from Monte Carlo simulation.
In each subfigure, $\rr_1$ and $\rr_2$ are held fixed, while the third
position $\rr_3$ is varied.  The test-particle position in each case
is $\rr_1$, which is on the left-hand side of the figure.  In the
color plot on the left, the first and second spheres are displayed in
grey.  On the bottom half of
figures~\ref{fig:triplet-contact-distribution}a and
\ref{fig:triplet-contact-distribution}c is the triplet distribution
function as computed using the approximation presented in this work.
The top half of those figures displays the triplet distribution
function from Monte Carlo.
Figures~\ref{fig:triplet-contact-distribution}b
and~\ref{fig:triplet-contact-distribution}d show the triplet
distribution function along the paths illustrated in the corresponding
2D plots.

Figures~\ref{fig:triplet-contact-distribution}a
and~\ref{fig:triplet-contact-distribution}b show the triplet
distribution
function when two of the positions $\rr_1$ and $\rr_2$ are in contact.
The triplet distribution function predicted by this work is slightly
asymmetric, indicating a failure in thermodynamic consistency.  Once
again, we observe the lack of oscillations at large distances.
Figures~\ref{fig:triplet-contact-distribution}c
and~\ref{fig:triplet-contact-distribution}d show the triplet distribution
function when two of the positions $\rr_1$ and $\rr_2$ are spaced so
that the third could just fit between them.  Again, the triplet
distribution function predicted by this work is slightly asymmetric,
and lacks the oscillations seen in the Monte Carlo simulations at
large distances.  In both cases, our agreement with simulation is
comparable to that of previous approximations.

\begin{figure}
  \begin{subfigure}{0.5\columnwidth}
    \includegraphics[width=\columnwidth]{figs/dadz-3-2.pdf}
    \vspace{-0.8cm}
    \caption{Sticky hard-sphere fluid}\label{fig:dadz-delta}
  \end{subfigure}
  \begin{subfigure}{0.5\columnwidth}
    \includegraphics[width=\columnwidth]{figs/dadz-square-well-3.pdf}
    \vspace{-0.8cm}
    \caption{Hard-core square well fluid}\label{fig:dadz-square-well}
  \end{subfigure}
  \begin{subfigure}{0.5\columnwidth}
    \includegraphics[width=\columnwidth]{figs/dadz-inverse-sixth-3.pdf}
    \vspace{-0.8cm}
    \caption{Hard-core inverse-sixth fluid}\label{fig:dadz-inverse-sixth}
  \end{subfigure}
  \caption{Plot of $\frac{da_1}{dz}$ near a hard wall.  (a) shows a
    sticky hard-sphere fluid defined by a pair potential
    $\delta(\sigma-r+\delta)$ where $\sigma$ is the hard-sphere
    diameter, and $\delta$ is an infinitesimal distance; (b) shows a
    square well fluid defined by a pair potential $\Theta(1.79
    \sigma-r)$; and (c) shows a hard-core inverse-sixth potential
    fluid with an attractive pair potential proportional to $r^{-6}$.
  }
  \label{fig:dadz}
\end{figure}

@@
